Flattering Advice: Avoiding Disappointment as a Driver of Gender Discrimination

### 2024-12-12 manuscript/manuscript.Rmd
Intro edits (in progress)

We frequently seek and offer advice, often with the hope of solving problems 
more effectively. Whether we are navigating personal dilemmas or making 
professional decisions, advice is seen as a valuable tool for improving 
outcomes. For instance, in some studies, participants are asked to estimate 
quantities or choose between lotteries [@Benjamin2015]. Advisors are often 
motivated to provide accurate advice [@Jonas2003], and engage in perspective-taking
to align with the preferences of the advisee [@Kray2000; @Kray1999]. When 
uncertain, they may even prefer that advisees rely less on their recommendations
[@Ache2020]. This suggests that people are generally concerned with offering 
good advice, assuming accuracy is the primary objective.

>>>>

Individuals seek (and give) advice on complex choices in their professional and
personal lives. While advice is widely viewed as a mechanism for improving 
outcomes (CITE), recent findings suggest systematic gender differences in 
the advice received: men are more likely to be encouraged toward competitive 
positions and risk-taking behaviors (CITE, if real), and women are less likely to receive
criticism (CITE). These differences in advice can result in gender gaps in
career advancement, as opportunities often depend on individuals applying (even
if they do not meet all requirements, CITE-women apply for jobs when they meet all reqs)
and direct feedback, even if unpleasant, can help people improve their performance.
In this paper, we propose and present evidence for a novel mechanism to explain
gendered patterns in advice received: advisers are motivated by interpersonal
considerations and do not want to disappoint the advisee.

Models of decision-making that incorporate belief utility (Loewenstein & Molnar) 
imply a tradeoff between the benefits from having accurate beliefs and the
hedonic costs of receiving bad news. For example, someone expecting a good outcome
may feel disappointment if the realized outcome is not as good as what they had
expected holding accurate beliefs and the hedonic
costs of receiving bad news, such as disappointment from a desired outcome
failing to materialize (Gul and Pesendorfer, I think, disappointment aversion, Golman et al 2017). 
Because outcomes are frequently evaluated with respect to expectations
(Koszegi and Rabin), advice that falls short of someone's expectation can be
painful. Thus, advice may be shaped by the *beliefs* about someone else's
expectations.

perpetuate gender gaps in career advancement, as many opportunities depend on 
individual initiative and self-selection. This paper proposes a novel mechanism 
to explain gendered patterns in advice-giving: advisors calibrate their 
recommendations to match perceived advisee expectations. In professional 
contexts where women typically demonstrate lower confidence levels, advisors 
may provide less ambitious guidance that aligns with these reduced expectations.
In two experimental studies, we establish evidence for this expectation-matching 
mechanism. Specifically, men are more confident in their performance on a task
than women and advisers, who observe performance and confidence but not gender,
give more flattering advice to men than to women. A third experiment shows that 
advisors strategically provide "flattering" advice to avoid disappointing their 
advisees, further illuminating how interpersonal dynamics shape advice content.

For instance, in some studies, participants are asked to estimate 
quantities or choose between lotteries [@Benjamin2015]. Advisors are often 
motivated to provide accurate advice [@Jonas2003], and engage in perspective-taking
to align with the preferences of the advisee [@Kray2000; @Kray1999]. When 
uncertain, they may even prefer that advisees rely less on their recommendations
[@Ache2020]. This suggests that people are generally concerned with offering 
good advice, assuming accuracy is the primary objective.
